---
title: "Three models for willingness to accept payment"
subtitle: "(chr, fct, num)"
execute:
  message: false
  warning: false
format: docx
editor_options: 
  chunk_output_type: console
---
```{r}

library(tidyverse)
library(chibukuplasticscollection)
```

## Data preparation

- Prepare wta variable which has a positive linear relationship with income (monthly_earn1)

```{r}

survey_model <- survey |> 
  select(id, age, gender, monthly_earn1) |> 
  # positive linear relationship for wta: with higher income, willingness to
  # accept payment increases
  mutate(wta = case_when(
    monthly_earn1 == "Less than MWK25,000" ~ 5,
    monthly_earn1 == "Between MWK25,000 to MWK49,999" ~ 10,
    monthly_earn1 == "Between MWK50,000 to MWK74,999" ~ 20,
    monthly_earn1 == "Between MWK75,000 to MWK99,999" ~ 40,
    monthly_earn1 == "Between MWK100,000 to MWK124,999" ~ 60,
    monthly_earn1 == "Between MWK125,000 to MWK149,999" ~ 90,
    monthly_earn1 == "Between MWK150,000 to MWK174,999" ~ 120,
    monthly_earn1 == "Between MWK175,000 to MWK200,000" ~ 160,
    monthly_earn1 == "More than MWK200,000" ~ 220
  ))

# write csv
survey_model |>
  write_csv(here::here("data/intermediate/survey_small_wta_model.csv"))
```

## Data transformation

- nearly 50% of responses cannot be used (I don't know, prefer not to say, NA)
- why so many NAs (20%) ?

```{r}
survey_model |> 
  count(monthly_earn1) |> 
  mutate(percent = n / sum(n) * 100) |> 
  knitr::kable()
```

```{r}
# define levels for factor
levels_monthly_earn <- c(
  "Less than MWK25,000",
  "Between MWK25,000 to MWK49,999",
  "Between MWK50,000 to MWK74,999",
  "Between MWK75,000 to MWK99,999",
  "Between MWK100,000 to MWK124,999",
  "Between MWK125,000 to MWK149,999",
  "Between MWK150,000 to MWK174,999",
  "Between MWK175,000 to MWK200,000",
  "More than MWK200,000"
)

survey_model_clean <- survey_model |> 
  filter(!monthly_earn1 %in% c("Prefer not to say", "I don't know")) |> 
  filter(!is.na(monthly_earn1)) |>
  rename(monthly_earn_chr = monthly_earn1) |> 
  mutate(monthly_earn_fct = fct_relevel(monthly_earn_chr, levels_monthly_earn)) |> 
  mutate(monthly_earn_num = case_when(
    monthly_earn_chr == "Less than MWK25,000" ~ 12500,
    monthly_earn_chr == "Between MWK25,000 to MWK49,999" ~ 37500,
    monthly_earn_chr == "Between MWK50,000 to MWK74,999" ~ 62500,
    monthly_earn_chr == "Between MWK75,000 to MWK99,999" ~ 87500,
    monthly_earn_chr == "Between MWK100,000 to MWK124,999" ~ 112500,
    monthly_earn_chr == "Between MWK125,000 to MWK149,999" ~ 137500,
    monthly_earn_chr == "Between MWK150,000 to MWK174,999" ~ 162500,
    monthly_earn_chr == "Between MWK175,000 to MWK200,000" ~ 187500,
    monthly_earn_chr == "More than MWK200,000" ~ 225000
  )) 
```

## Data visualisation & modeling

### Income not ordered (i.e. alphabetically)

```{r}

survey_model_clean |> 
  ggplot(aes(x = monthly_earn_chr, y = wta)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))

```

```{r}
model_chr_fit <- lm(wta ~ monthly_earn_chr, data = survey_model_clean)
```

**Model summary**

```{r}
summary(model_chr_fit)
```

**Model coefficients**

```{r}
broom::tidy(model_chr_fit)
```


### Income ordered 

```{r}
survey_model_clean |> 
  ggplot(aes(x = monthly_earn_fct, y = wta)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90))

```

```{r}
model_fct_fit <- lm(wta ~ monthly_earn_fct, data = survey_model_clean)
```

**Model summary**

```{r}
summary(model_fct_fit)
```

**Model coefficients**

```{r}
broom::tidy(model_fct_fit)
```

### Income as numeric

```{r}
survey_model_clean |> 
  ggplot(aes(x = monthly_earn_num, y = wta)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  theme(axis.text.x = element_text(angle = 90))
```


```{r}
model_num_fit <- lm(wta ~ monthly_earn_num, data = survey_model_clean)
```

**Model summary**

```{r}
summary(model_num_fit)
```

**Model coefficients**

```{r}
broom::tidy(model_num_fit)
```


